Search for WIMP Inelastic Scattering Off Xenon Nuclei With XENON100 Data

### 2017-01-19 Inelastic/analysis.tex
continued_with_section3_A_B_C_some_comments_in_red

This analysis is performed using XENON100 run\,II science data, which corresponds to a data set with an exposure of 224.6~live days. The detector response to
ERs has been characterised with $^{60}$Co and $^{232}$Th calibration sources, while the response to NRs was calibrated with an $^{241}$AmBe $(\alpha,n)$-source. This fast neutron source gives rise to elastic and inelastic neutron-nucleus scatters, and can thus be employed to define the expected signal region.

@@

A particle interaction in the liquid xenon produces an S1 and a correlated S2 signal with a certain number of photoelectrons (PE) observed by the PMTs. The non-uniform scintillation light collection by the PMT arrays, due to solid angle effects, Rayleigh scattering length, reflectivity, transmission of the electrodes, etc, lead to a position-dependent S1 signal. The warping of the top meshes (inducing a variation in the width of the gas gap between the anode and the liquid-gas interface), the absorption of electrons by residual impurities, as they drift towards the gas region, as well as solid angle effects lead to a position-dependent S2 signal. These signals are thus corrected in 3 dimensions, using various calibration data~\cite{Aprile:2011dd}, with the corrected quantities denoted as cS1 and cS2.

\subsection{Signal Region}

@@

Other than falling in the defined region of interest, events are required to fulfil several selection criteria, which can be summarised as follows:
selection aimed to reduce noise impact and including energy (S1) and S2 thresholds, veto of events with energy release in the detector's outer volume, 
and additionally, events must be of single scatter nature and fall into a predefined fiducial volume. This analysis follows the 
selection criteria described in detail in \cite{dataAnalysis} for Run-II, with only few exceptions reported in what follows. 
%only few modification have been designed specifically for this analysis and discussed below. 
The selection on S2 width as a function of drift time has been optimized on a sample of events selected from the 39.6~keV line
and set to a 95\% acceptance on these. Events are required to be single scatter by applying a threshold on the 
second largest S2 peak size,  for this analysis this threshold has been optimized to 160~PE and set constant as function of S2 signal size. 
Finally the chosen fiducial volume corresponds to 34~Kg of liquid xenon.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{images/bkg_in_sr.png}
  \caption{Signal region and control region.}
  \label{fig:SR}
\end{figure}

\subsection{Signal Simulation} 
%description of the simulated signal, few words about cross checks MC matching.

The detector response to inelastic scattering of WIMPs  off $^{129}$Xe nucleus was simulated using an empirical model.
The total deposited energy is divided into two independent contributions: one related to the 39.6~keV de-excitation photon and the other  relative to
the simultaneous nuclear recoil of the xenon atom. The number of photons and charge yield detected is simulated separately for each contribution
and then added together.

The distribution in the cS1-cS2 plane of an ER induced by the de-excitation photon is simulated assuming a two dimensional normal pdf, $f(cS1,cS2)$, 
described except a normalization factor in equation~\ref{f:2dgaus}:
\begin{multline}
	f(cS1,cS2)  = exp \Big( -\frac{1}{2(1-\rho^2)} \Big[ \frac{(cS1 - \mu_{cS1})^2}{\sigma_{cs1}^2} + \\ 
	 \frac{(cS2 - \mu_{cS2})^2}{\sigma_{cs2}^2} - \frac{2\rho(cS1 - \mu_{cs1}) (cS2 - \mu_{cs2})} {\sigma_{cs1}\sigma_{cs2}} \Big] \Big) 
	%f(cS1,cS2) ~ = ~  \frac{1}{2 \pi \sigma_{cs1} \sigma_cs2 \sqrt{1-\rho^{2}} } exp \Big( -\frac{1}{2(1-\rho^2)} \Big[ \frac{(cS1 - \hat{cS1})^2}{\sigma_{cs1}^2} \Big] \Big)
\label{f:2dgaus}
\end{multline}
where $\mu_{cS1}$ and $\mu_{cS2}$ 
represents the average observed cS1 and cS2 given a 39.6~keVee , $\sigma_{cs1}$ and $\sigma_{cs2}$ are the standard deviation in cS1 and cS2 respectively,
while $\rho$ stands for the correlation between cS1 and cS2.
The detector related light yield $L_y$  at 39.6~keV, necessary to evaluate the average number of photon detected ($\mu_{cS1}$),
%with AmBe calibration has been found to have a large systematic uncertainty
%due to the contribution of the nuclear recoil, so the light yield used to describe the number of photon detected 
is obtained as the result of a NEST model~\cite{NEST,Geant1,Geant2} fit to data collected with several lines.  The same model is used to predict the charge yield at
39.6~keV which is then scaled according to the detector's secondary scintillation gain $Y$, determined from detector response to single electrons~\cite{SingleE}.
The detector resolution at 39.6~keV in cS1 and cS2 has been measured to be respectively 15.8\% and 14.7\% and used to extract the standard 
deviations $\sigma_{cs1}$, $\sigma_{cs2}$.  The correlation parameter is assumed to be independent of energy (at least in the considered range) and measured
using the 164~keV xenon activated line by $^{124}$AmBe calibration data, this line is chosen since allows to disentangle efficiently contribution from nuclear recoil.
The measured correlation is $\rho \, = \, -0.45 \pm 0.10$.

The cS1 and cS2 distributions from NR contribution are predicted starting from the nuclear recoil energy spectrum
of WIMPs inelastic interaction~\cite{inelastic_th}. The average cS1 and cS2 are given by equations~\ref{f:cs1} and~\ref{f:cs2} respectively,
where $\mathcal{L}_{eff}$ is the liquid xenon NR relative scintillation efficiency, while $S_{ee} \, = \, 0.58$  and $S_{nr} \, = \, 0.95$ describe the scintillation 
quenching due to the electric field \cite{ScintQuenching}. The parameterization and uncertainties of $\mathcal{L}_{eff}$ as a function of $E_{nr}$ are based on existing 
direct measurements \cite{run8Result}. The light yield at 122~keVee originate from the same NEST model fit as described above. For the cS2 the parameterization 
of $Q_{Y}(E_{nr})$ is taken from \cite{QY}.
Finally all detector related resolution effects are introduced following the prescriptions described in \cite{dataAnalysis}.

@@

The pdf of the ER and NR contributions are then convoluted together to obtain the overall pdf of the signal.
%The two contribution are summed together in an event-by-event basis using random extraction from the two pdfs. 
A 2D (cS1 versus cS2) acceptance map is applied to the signal pdf to reproduce data selection effects. Acceptances are computed separately for each selection 
criteria using $^{124}$AmBe calibration sample, selections as the outer volume veto and the single scatter interaction represent an exception  and 
a dedicated computation has been performed in these cases. The selections acceptance average in the region of interest to about $0.80 \pm 0.05$. 
Figure~\ref{fig:SR} shows an example of full simulated signal model for a WIMP of 100~GeV mass.

The signal simulation procedure has been validated reproducing the 39.6~keV xenon line from interaction due to
$^{124}$AmBe source and has been compared to data. For this comparison the proper  $^{124}$AmBe nuclear recoil and acceptances
were simulated. The simulated events were found in agreement  with calibration data within statistical uncertainties.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{images/wimp_in_sr.png}
  \caption{Signal region and control region, for WIMP of mass 100~GeV.}
  \label{fig:SR2}
\end{figure}

>>>>

This analysis is performed using XENON100 Run-II science data, which corresponds to a data set with an exposure of 224.6~live days. The detector response to
ERs has been characterised with $^{60}$Co and $^{232}$Th calibration sources, while the response to NRs was calibrated with an $^{241}$AmBe $(\alpha,n)$-source. This fast neutron source gives rise to elastic and inelastic neutron-nucleus scatters, and can thus be employed to define the expected signal region for inelastic WIMP-nucleus scatters.

@@

A particle interaction in the liquid xenon produces an S1 and a correlated S2 signal with a certain number of photoelectrons (PE) observed by the PMTs. The non-uniform scintillation light collection by the PMT arrays, due to solid angle effects, Rayleigh scattering length, reflectivity, transmission of the electrodes, etc, lead to a position-dependent S1 signal. The warping of the top meshes (inducing a variation in the width of the gas gap between the anode and the liquid-gas interface), the absorption of electrons by residual impurities as they drift towards the gas region, as well as solid angle effects lead to a position-dependent S2 signal. These signals are thus corrected in 3 dimensions, using various calibration data, as detailed in~\cite{Aprile:2011dd,Aprile:2012vw}, with the corrected quantities denoted as cS1 and cS2, and defined in~\cite{Aprile:2012vw}. The trigger efficiency in thus run was 100\% for S2$>$300\,PE.

\subsection{Signal Region and Event Selection}

@@

Apart from the condition to occur in the defined region of interest in the cS2-cS1-plane, valid events are required to fulfil several selection criteria, 
which can be summarised as follows: data quality cuts, energy selection and S2 threshold cut, veto cut for events with energy release in the detectors active LXe shield, 
selection of single scatter events and of a predefined fiducial volume.

This analysis follows the selection criteria described in detail in \cite{Aprile:2012vw} for Run-II, with only a few exceptions reported in what follows. 
The cut on the S2 signal width as a function of drift time (where the maximal drift time is 176\,$\mu$s and the width values range from $\sim$1-2\,$\mu$s) has been optimised on a sample of events selected from the 39.6\,keV line and set to a 95\% acceptance on these. Events are required to be single-scatters by applying a threshold cut on the size of the 
second largest S2 peak. For this analysis, the threshold has been optimised to 160\,PE and set constant as function of S2 signal size. 
Finally, the chosen fiducial volume corresponds to 34\,kg of liquid xenon.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{images/bkg_in_sr.png}
  \caption{Signal (1-9) and control (CR1 and CR2) regions for the inelastic WIMP-$^{129}$Xe interaction in the cS2 versus cS1 plane. {\textcolor{red} {I think we must explains the shape of these regions.}}}
  \label{fig:SR}
\end{figure}

\subsection{Signal Simulation}

The detector response to inelastic WIMP-$^{129}$Xe interactions was simulated using an empirical signal model.
The total deposited energy is divided into two independent contributions: one coming from the 39.6\,keV de-excitation photon and the other  from  
the simultaneous nuclear recoil of the xenon atom. The detected light (S1) and charge (S2) signals are simulated separately for each of the two contributions 
and then added together. This is due to the fact that the light and charge yields depend on the type of interaction (ER vs. NR), and on the deposited energy.

The distribution of an ER induced by the de-excitation photon in the cS1-cS2 plane  is simulated assuming a two dimensional normal probability distribution function (pdf), $f(cS1,cS2)$, 
described (apart from a constant normalisation factor) by the following equation:

\begin{multline}
	f(cS1,cS2)  = {\rm exp} \Big\{ -\frac{1}{2(1-\rho^2)} \Big[ \frac{(cS1 - \mu_{cS1})^2}{\sigma_{cs1}^2} + \\ 
	 \frac{(cS2 - \mu_{cS2})^2}{\sigma_{cs2}^2} - \frac{2\rho\cdot(cS1 - \mu_{cs1}) (cS2 - \mu_{cs2})} {\sigma_{cs1}\sigma_{cs2}} \Big] \Big\}
	%f(cS1,cS2) ~ = ~  \frac{1}{2 \pi \sigma_{cs1} \sigma_cs2 \sqrt{1-\rho^{2}} } exp \Big( -\frac{1}{2(1-\rho^2)} \Big[ \frac{(cS1 - \hat{cS1})^2}{\sigma_{cs1}^2} \Big] \Big)
\label{f:2dgaus}
\end{multline}
where $\mu_{cS1}$ and $\mu_{cS2}$ 
represent the average observed cS1 and cS2 signals given a 39.6\,keV ER, $\sigma_{cs1}$ and $\sigma_{cs2}$ are the standard deviation in cS1 and cS2 respectively,
while $\rho$ stands for the correlation between the cS1 and cS2 signals. The detector-related light yield $L_y$  at 39.6\,keV, necessary to evaluate the average number of prompt photons detected 
($\mu_{cS1}$), is obtained from the NEST model~\cite{NEST,Geant1,Geant2} fit to data collected with several $\gamma$-lines.  The average light yield at 39.6\,keV is 2.7\,PE/keV.

The same model is used to predict the charge yield at 39.6\,keV, which is afterwards scaled according to the detector's secondary scintillation gain $Y$. 
 The latter is determined from detector's response to single electrons~\cite{SingleE}.
The energy resolution at 39.6\,keV in cS1 and cS2 has been measured to be 15.8\% and 14.7\%, respectively, and is used to extract the standard 
deviations $\sigma_{cs1}$, $\sigma_{cs2}$.  The correlation parameter is assumed to be independent of energy (at least in the considered narrow energy range) and measured
using the 164\,keV line from the decay of the $^{131m}$Xe isomer ($T_{1/2}$=11.8\,d) produced during the  $^{124}$AmBe run. This $\gamma$-line is chosen because it allows to disentangle efficiently the contribution from the nuclear recoil. The measured correlation coefficient is $\rho \, = \, -0.45 \pm 0.10$.

The cS1 and cS2 distributions from the NR contribution are predicted starting from the expected nuclear recoil energy spectrum
of WIMP inelastic interactions~\cite{Baudis:2013bba}. The average cS1 and cS2 are given by equations~\ref{f:cs1} and~\ref{f:cs2} respectively,
where $\mathcal{L}_{eff}$ is the liquid xenon relative scintillation efficiency for NRs, while $S_{ee} \, = \, 0.58$  and $S_{nr} \, = \, 0.95$ describe the scintillation 
quenching  of ER and NRs, respectively, due to the electric field \cite{ScintQuenching}. The parameterisation and uncertainties of $\mathcal{L}_{eff}$ as a function of nuclear recoil energy $E_{nr}$ are based on existing 
direct measurements \cite{run8Result}. The light yield for 122\,keV ERs is taken from the same NEST model fit as described above. For cS2, the parameterisation 
of $Q_{Y}(E_{nr})$ is taken from \cite{QY}. Finally, all detector related resolution effects are introduced following the prescriptions described in \cite{Aprile:2012vw}.

@@

The pdf of the ER and NR contributions are then convoluted together to obtain the overall pdf of the expected signal.
A 2D (cS1 versus cS2) acceptance map is applied to the signal pdf to reproduce data selection effects. Acceptances are computed separately for each selection 
criteria using the $^{124}$AmBe calibration sample. Other selections such as the liquid xenon veto cut, and the single-scatter interaction represent an exception  and 
a dedicated computation has been performed in these cases. The combined acceptance  of all selection criteria in the region of interest averages to $\sim$$(0.80\pm0.05)$. 
Figure~\ref{fig:SR} shows an example of fully simulated signal model for a WIMP mass of 100\,GeV/c$^2$.

The signal simulation procedure has been validated by reproducing the 39.6\,keV xenon line from interactions due to neutrons from the 
$^{124}$AmBe source and comparison to data. For this comparison, the proper  $^{124}$AmBe nuclear recoil and acceptances
were simulated. The simulated events were in agreement  with calibration data within statistical uncertainties. 
{\textcolor{red} {Should we show an example, namely a figure from the signal model note?}}

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{images/wimp_in_sr.png}
  \caption{Simulated signal (1-9) and control (CR1 and CR2) regions for a WIMP mass 100\,GeV/c$^2$.}
  \label{fig:SR2}
\end{figure}
